---
title: 'Springboard: Foundations of Data Science'
author: "Tom Skawski II"
date: "July 24, 2016"
output: html_document
subtitle: Data Wrangling Exercise 2
---

```{r setup, include=FALSE}
# no idea what this does or why it's here, but I'm leaving it.
# found this in the knitr help
knitr::opts_chunk$set(echo = TRUE)
```

### Step 00 - Prepare Directory and load packages
```{r, message=FALSE}
# Mac:
setwd("~tomskawski/Dropbox/Data Science/Data Wrangling/DW2")
# PC
# setwd("C:/Users/tskawski.WESTERN/Desktop/Dropbox/Data Science/Data Wrangling/DW2")

# install.packages("tidyr")
library("tidyr")

# install.packages("dplyr")
library("dplyr")
```

### Step 0: Load the data in RStudio
Save the data set as a CSV file called titanic_original.csv and load it in RStudio into a data frame.
```{r}
mydata <- read.csv("titanic_original.csv", stringsAsFactors = FALSE)
```

```{r}
# Check class of columns. Should not be factors.
sapply(mydata, class)
```

### Step 1 - Port of embarkation
The embarked column has some missing values, which are known to correspond to passengers who actually embarked at Southampton. Find the missing values and replace them with S. (Caution: Sometimes a missing value might be read into R as a blank or empty string.)

```{r}
mydata %>% group_by(embarked) %>% summarise(name = n())
```

```{r}
# REGEX hunting yet again
grep("^$", mydata$embarked)
```

```{r}
mydata$embarked <- gsub("^$", "S", mydata$embarked)
```

```{r}
# Re-check
mydata %>% group_by(embarked) %>% summarise(name = n())
```

### Step 2: Age
You’ll notice that a lot of the values in the Age column are missing. While there are many ways to fill these missing values, using the mean or median of the rest of the values is quite common in such cases.

Calculate the mean of the Age column and use that value to populate the missing values

Think about other ways you could have populated the missing values in the age column. Why would you pick any of those over the mean (or not)?

```{r}
# tried to check by age, but too many fields (how would I parse more?)
mydata %>% group_by(age) %>% summarise(name = n())
```

Found solution [here](https://stackoverflow.com/questions/25835643/replacing-missing-values-in-r-with-column-mean)

```{r}
mydata$age[is.na(mydata$age)] <- mean(mydata$age, na.rm = TRUE)
# mydata$age will show all the data
```

### Step 3: Lifeboat
You’re interested in looking at the distribution of passengers in different lifeboats, but as we know, many passengers did not make it to a boat :-( This means that there are a lot of missing values in the boat column. Fill these empty slots with a dummy value e.g. the string 'None' or 'NA'

```{r}
mydata$boat <- gsub("^$", "None", mydata$boat)
```

### Step 4: Cabin
You notice that many passengers don’t have a cabin number associated with them.
Does it make sense to fill missing cabin numbers with a value?
What does a missing value here mean?
You have a hunch that the fact that the cabin number is missing might be a useful indicator of survival. Create a new column has_cabin_number which has 1 if there is a cabin number, and 0 otherwise.

```{r}
# regex doesn't play nice between grep and as functions (using .+ for grep and != "" for as.integer)
head(grep(".+", mydata$cabin, value = TRUE), 10)

mydata$has_cabin_number <- as.integer(mydata$cabin != "")
```

### Step 6: Submit the project on Github
Include your code, the original data as a CSV file titanic_original.csv, and the cleaned up data as a CSV file called titanic_clean.csv.

```{r}
# Make file and submit
write.csv(mydata, file = "titanic_clean.csv")

# Used github desktop, drag and drop
```